---
title: "p8105_hw1_ms5533"
author: "Marisa Sobel"
date: 2018-09-21
output: github_document
---

# Set-Up

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

### Create a data frame comprised of:
* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r df_create}

set.seed(1)

prob1_df = tibble(
  list = sample(0:5, 10, replace = TRUE), 
  greater_2 = list > 2, 
  vec_car = c("Will", "I", "ever", "be", "able", "to", "remember", "any", "R", "commands"), 
  vec_factor = c("yes", "no", "yes", "no", "yes", "no", "yes", "no", "yes", "no")
)
```

### Take the mean of each variable in your dataframe. What works and what doesn’t? Why?

```{r means}
mean(prob1_df$list)
mean(prob1_df$greater_2)
mean(prob1_df$vec_car)
mean(prob1_df$vec_factor)
```

The mean of "list"" and the logical vector "greater_2" are possible. "list" is a list of numbers, and therefore a mean is possible and equals `r mean(prob1_df$list)`. "greater_2" is a logical vector with a dichotomized outcome (TRUE/FALSE), coded as 1 or 0, and equals `r mean(prob1_df$greater_2)`. Means of "vec_car" and "vec_factor" cannot compute as they are not numeric. 

### Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens?

```{r means_as_numeric, eval=FALSE}
mean(as.numeric(prob1_df$greater_2))
mean(as.numeric(prob1_df$vec_car))
mean(as.numeric(prob1_df$vec_factor))
```

When applying the "as.numeric" function to the logical vector, it produced the same result, `r mean(as.numeric(prob1_df$greater_2))`. The "as.numeric" function did not change the outcome of the character or factor vectors. 

### Convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?

```{r converting, eval=FALSE}
as.numeric(as.factor(prob1_df$vec_car))
as.numeric(as.character(prob1_df$vec_factor))
```

Converting the character vector to factor and then to numeric is possible; it returns a numeric list. If first orders the words alphabetically, then it applies numbers to the alphabetical list. The output then is the corresponting number to the word in the original order. 

Converting the factor vector to character to numeric, however, is not possible. 

# Problem 2

```{r}

```
