---
title: "p8105_hw1_ms5533"
author: "Marisa Sobel"
date: 2018-09-21
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

```{r df1_create}
set.seed(1)
prob1_df = tibble(
  list = sample(0:5, 10, replace = TRUE), 
  greater_2 = list > 2, 
  vec_car = c("Will", "I", "ever", "be", "able", "to", "remember", "any", "R", "commands"), 
  vec_factor = c("yes", "no", "yes", "no", "yes", "no", "yes", "no", "yes", "no")
)
```

### Mean of each variable in dataframe

```{r means}
mean(prob1_df$list)
mean(prob1_df$greater_2)
mean(prob1_df$vec_car)
mean(prob1_df$vec_factor)
```

The mean of _**list**_ and the logical vector _**greater_2**_ are possible. _**list**_ is a list of numbers, and therefore a mean is possible and equals `r mean(prob1_df$list)`. _**greater_2**_ is a logical vector with a dichotomized outcome (TRUE/FALSE), coded as 1 or 0, and equals `r mean(prob1_df$greater_2)`. Means of _**vec_car**_ and _**vec_factor**_ cannot compute as they are not numeric. 

### Means as numeric variables

```{r means_as_numeric, eval=FALSE}
mean(as.numeric(prob1_df$greater_2))
mean(as.numeric(prob1_df$vec_car))
mean(as.numeric(prob1_df$vec_factor))
```

When applying the _**as.numeric**_ function to the logical vector, it produced the same result, `r mean(as.numeric(prob1_df$greater_2))`. The _**as.numeric**_ function did not change the outcome of the character or factor vectors. 

### Changing classes of variables

```{r converting, eval=FALSE}
as.numeric(as.factor(prob1_df$vec_car))
as.numeric(as.character(prob1_df$vec_factor))
```

Converting the character vector to factor and then to numeric is possible; it returns a numeric list. If first orders the words alphabetically, then it applies numbers to the alphabetical list. The output then is the corresponting number to the word in the original order. 

Converting the factor vector to character to numeric, however, is not possible. 

# Problem 2

```{r df2_create}
set.seed(2)
prob2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  greater_0 = x + y > 0, 
  TF_num = as.numeric(greater_0),
  TF_fac = as.factor(TF_num)
)
```
```{r df2_str, include=FALSE}
str(prob2_df)
```

The data frame created has a size of `r nrow(prob2_df)` data points. The mean of _**x**_ is equal to `r round(mean(prob2_df$x), digits = 5)` and the median of _**x**_ is equal to `r round(median(prob2_df$x), digits = 5)`.  `r sum(prob2_df$TF_num)` data points of `r nrow(prob2_df)` total are above 0 (`r sum(prob2_df$TF_num)`/`r nrow(prob2_df)`). 

### Scatterplot of _y_ vs _x_ by logical variable

```{r yx_scatter_log, echo=FALSE}
ggplot(prob2_df, aes(x = x, y = y, color = greater_0)) + geom_point()
```
```{r yx_scatter_log_save, include=FALSE}
ggsave
```

There is a clear distribution of values of _**x + y > 0**_ and _**x + y < 0**_ along the line _**y = -x**_. 

### Scatterplot of _y_ vs _x_ by numeric variable

```{r yx_scatter_num, echo=FALSE}
ggplot(prob2_df, aes(x = x, y = y, color = TF_num)) + geom_point()
```

The same distribution is seen as above between points with values of _**x + y > 0**_ and _**x + y < 0**_ along the line _**y = -x**_. However, as they are now color-coded by a numeric variable and not a categorical variable, the legend shows a ramp of colors possible based on the value. Since it is a dichotomous variable, there are only two options, 1 and 0. Therefore the chart looks the same. 

### Scatterplot of _y_ vs _x_ by factor variable

```{r yx_scatter_fac, echo=FALSE}
ggplot(prob2_df, aes(x = x, y = y, color = TF_fac)) + geom_point()
```

The same distribution is seen as in the first plot, as TRUE is coded as 1 and FALSE is coded as 0. 




